---
title: |
  | A Sudden Surge in Climate Awareness:
  | Evidence from German Political Polling in 2019
title-meta: ""
subtitle: | 
  | *Master Thesis Research Hypothesis*
author: | 
  | Johann-Friedrich Salzmann [\Letter](mailto:mail@jfsalzmann.com)
author-meta: Johann-Friedrich Salzmann
date: "last edited Feb 07, 2023"
output:
  pdf_document:
    number_sections: false
    citation_package: natbib
    keep_tex: true
    df_print: kable
    latex_engine: pdflatex
    fig_caption: true
bibliography:
  - ../Literatur/bibtex/library.bib
  - packages.bib
urlcolor: MidnightBlue
linkcolor: MidnightBlue
citecolor: MidnightBlue
geometry: "left=1in,right=1in,top=1in,bottom=1in"
fontsize: 12pt
biblio-style: agsm
# dcu apacite
header-includes:
  - \usepackage{mathtools}
  - \usepackage{amsmath}
  - \numberwithin{equation}{section}
  - \numberwithin{table}{section}
  - \numberwithin{figure}{section}
  - \usepackage{float}
  - \usepackage{gensymb}
  - \usepackage{setspace}\singlespacing
  - \usepackage{titlesec}
  - \titlespacing*{\section}{0pt}{2cm}{0.5cm}
  - \titlespacing*{\subsection}{0pt}{1cm}{1cm}
#  - \usepackage[ngerman]{babel}
  - \usepackage[nottoc]{tocbibind}
#  - \setcounter{secnumdepth}{2}
  - \setcounter{tocdepth}{2}
  - \widowpenalties 3 10000 10000 150
  - \displaywidowpenalties 3 10000 10000 150
  - \clubpenalties 3 10000 10000 150
  - \interfootnotelinepenalty 5000
  - \raggedbottom
  - \usepackage{enumitem}
  - \setlist{itemsep=3pt,topsep=5pt,parsep=0pt,partopsep=0pt}
  - \usepackage{marvosym}
  - \usepackage{hyperref}
  - \hypersetup{bookmarksnumbered=true}
  - \usepackage{bookmark}
#  - \newcommand{\sectionbreak}{\clearpage\phantomsection}
  - \let\Begin\begin
  - \let\End\end
  - \newcommand{\Newrow}{\\}
  - \addtolength{\skip\footins}{2pc plus 5pt}
#  - \AtBeginDocument{\pdfbookmark[section]{Deckblatt}{front}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results='hide', message=FALSE, warning=FALSE)
knitr::opts_chunk$set(fig.height = 4, fig.width = 6)
options(knitr.kable.NA = '',knitr.table.format = "latex")
#knitr::read_chunk('ba.R')
library(kableExtra)
library(readxl)
library(dplyr)
```

\renewcommand{\harvardand}{\&}
\setcitestyle{authoryear,round,aysep={},notesep={: }}
\nocite{Nowak1990,Quattrociocchi2014,Mason2007,Lavine1996,Castellano2009,Winkelmann2022,Xia2011,GRONLUND2005,Friedkin2011,Zhu2022,Ben-Naim2005,Otto2020a}

```{r bibprepare}
knitr::write_bib(c(
  loadedNamespaces(), 'knitr', 'rmarkdown', 'tidyverse', 'kableExtra'
), 'packages.bib')
```

My Master Thesis project is motivated by my aim to understand **social change dynamics** with a special focus to Germany's rapid change in public opinion on climate policy and its importance in 2019.

Within a few months in 2019, a major macro-level shift in public opinion could be observed [@ForschungsgruppeWahlen2023, MIP data] with regards to the perceived severity, urgency and relevancy of climate change. This sudden change may have been caused by an intensified public debate triggered by the global climate movement.


## Hypotheses

1. A variety of indicators prove that there actually was an unusual surge in climate change awareness, beyond usual/expected fluctuation (reflected through political opinion polls, most important issue surveys, etc.).
2. That surge can be observed across political parties and socio-economic groups (not necessarily equally strong, but significant).
3. The change in perception/political preferences significantly influenced party preferences accordingly.


## Data

The political polling data sets listed below are available through GESIS^[GESIS resources: https://www.gesis.org/wahlen/wahlen-home]. I would like to analyse different data sets and gather descriptive, explorative, and inferential insights.

### GLES Panel

- @ZA6838, free of charge
- longitudinal panel, twice a year, ~10k participants
- Panel 2016-2021 (waves 1-15), particular interest: May, Nov 2019 waves
- MIP, topic detail questions, party preferences

### ZDF Politbarometer

- @ZA7599, free of charge
- longitudinal trend/cross-section, bi-weekly, ~1.25k participants
- particular interest: Cumulated Data Set 2019 (more available)
- party preferences, politicians' popularity, current policy topics


### ARD Deutschlandtrend

- @ZA7755, 30 EUR for each year of data
- longitudinal trend/cross-section, monthly, ~1-1.5k participants
- particular interest: 2019 edition (more available)
- MIP, party preferences, politicians' popularity, current policy topics


### Forsa-Bus

- @ZA6850, 30 EUR for each year of data
- longitudinal trend/cross-section, daily, ~500 participants
- particular interest: 2019 edition (more available)
- party preferences, politicians' popularity



\footnotesize